A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility

### 2019-06-13 paper.Rmd
Pushed draft of conclusions

TBD

>>>>

# Conclusions

The open science literature has been advocating the importance of data sharing for a long time. However many of these articles appeal to the broader benefits of data sharing but rarely talk about the practical considerations of making data available alongside code. Trivial reproducibility has always relied up on the idea that as long as code exists on a platform such as GitHub and clearly lists open source dependencies, one should be able to run the code at a future time. Containerization efforts such as Docker have made it even easier to capture system dependencies and rapidly launch containers. But true reproducibility requires not just code but also data.

Over the years researchers have made data available in a multitude of unreliable ways, including offering downloads from university websites, FTP repositories, and password protected servers, all of which are prone to bit rot. In rare cases where the data are deposited in a permanent archive, insufficient metadata and missing data processing scripts have made it harder and time intensive to map these to code.  Our goal here is to describe a variety of ways in which a researcher can ship data, analogous to how easy it has become to push code to services like GitHub. The choice of technology and method all depends on the nature of the data, size, and complexity. Once the issue of access is solved, we also discuss how much metadata to include in order to make the data useful. Ultimately there is a tradeoff to consider on a case by case basis. Long-term studies with complex protocols require structured metadata which more standardized datasets such as those coming from sensors require far less effort.

The key message for the reader is that accessing data for reproducibility should come with minimal barriers. Having users jump through data usage forms or other barriers is enough of a roadblock to dissuade users, which over time will make it hard to track the provenance of critical studies. For data with broad impact, one should further invest effort in documenting the metadata (either unstructured or structured depending on the complexity) and also focus on ensuring that at least one archival copy exists in a permanent archive. Following this continuum can ensure that more and more computational work becomes readily reproducible without unreasonable effort.
